---
title: 'Data management, plotting and statistical testing'
subtitle: 'Practice Project in R course'
author: 'Yufan Yao'
date: '`r Sys.Date()`'
output: 
  html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
---
# Introduction
This document is 

# Setup environment and data import
## Load libraries
```{r}

```

## Set working directory
```{r}

```

## Import the datasets
```{r}

```

## Visually inspect imported data.frames
```{r}

```

# Data Management
## Change variables
### Set variables names
```{r}

```

### Transform dataset from wide to long format
```{r}

```

### Format Numeric variables
```{r}

```

### Formate categorical variables
```{r}

```

### Create new variables
```{r}

```

## Merge datasets
```{r}

```

## Re-arrange the data
### Reorder variables
```{r}

```

### Reorder rows
```{r}

```

# Numerical and graphical data summary
## Numerical summary of time-invariant variables
```{r}

```

## Graphical summary of time-invariant variables
### Histograms
```{r}

```

### Correlation matrix
```{r}

```

### Box plots
```{r}

```

## Numerical summary of time-variant variables
```{r}

```

## Graphical summary of time-variant variables
### Scatterplots
#### Spaghetti plots
```{r}

```

#### Panel plots
```{r}

```

# Statistical testing
```{r}

```

