---
title: 'Data management, plotting and statistical testing'
subtitle: 'Practice Project in R course'
author: 'Yufan Yao'
date: '`r Sys.Date()`'
output: 
  html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: true
---
# Introduction
This document is 

# Setup environment and data import
## Load libraries
```{r load packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

## Set working directory
```{r wd}
getwd()
```

## Import the datasets
```{r import dataset}
# read.csv(), read.table()
data_pd <- read.csv("BPI889_PD_run-in.csv", header = TRUE)
data_cov <- read.table("BPI889_demographics.tab", header = TRUE, sep = "")
```

## Visually inspect imported data.frames
```{r check}
str(data_pd)
str(data_cov)
```

# Data Management
## Change variables
### Set variables names
```{r}
names(data_pd)[1] <- "ID" # change the name of column in dataframe using names()
```

### Transform dataset from wide to long format
```{r long format}
# gather() to transform wide to long format
tidy_pd <- gather(
  data_pd,
  key = "TIME",
  value = "CD4",
  -ID) # exclude ID column
```

### Format Numeric variables
```{r}
# check the class of the variable
summary(tidy_pd)
# CD4 --> as.numeric()
tidy_pd$CD4 <- as.numeric(tidy_pd$CD4)
# TIME, only keeps numbers/ remove part of the string using gsub()
tidy_pd$TIME <- gsub("day.", "", tidy_pd$TIME)
tidy_pd$TIME <- as.numeric(tidy_pd$TIME)
```

### Format categorical variables
```{r}
# check the class of the variables
summary(data_cov)
# factorize SEX, and rename F/M to Female and Male, using factor()
data_cov$SEX <- factor(data_cov$SEX ,
                       levels = c("F","M"),
                       labels = c("Female", "Male"))
# factorize TB, rename 0/1 to HIV/HIV/TB using factor()
data_cov$TB <- factor(data_cov$TB,
                      levels = c(0, 1),
                      labels = c("HIV", "HIV+TB"))
```

### Create new variables
```{r}
# calculate BMI, BMI = weight/height2, weight - kg, height - m
# change unit of HT to m
data_cov$HT <- data_cov$HT/100
bmi <- function(WT, HT){
  bmi <- WT/(HT^2)
  bmi <- round(bmi, digits = 1) # limits decimals into 1 decimal using round()
  return(bmi)
}
data_cov$BMI <- bmi(data_cov$WT, data_cov$HT)
data_cov$CBMI <- NA
# ifelse() to assign values 0,1,2,3 to the variable CBMI
data_cov$CBMI <- ifelse(data_cov$SEX == "Male" & data_cov$BMI <= 20 | data_cov$SEX == "Female" & data_cov$BMI <= 18, 0,
                        ifelse(data_cov$BMI <= 25, 1, 
                               ifelse(data_cov$BMI <= 30, 2, 3)))
# re-class 0,1,2,3 to underweight, lean, overweight, obese with labels using factor()
data_cov$CBMI <- factor(data_cov$CBMI,
      levels = c(0,1,2,3),
      labels = c("underwight", "lean", "overweight", "obese"))
# summary of CBMI, to see the amount of each factor
table(data_cov$CBMI)
```

## Merge datasets
```{r}
# use three ways to merge datasets to see the differences
# cbind()
data_cbind <- cbind(tidy_pd, data_cov)
# merge(x,y, all=TRUE), full outer join
data_merge <- merge(tidy_pd, data_cov, all = TRUE)
# merge(x,y,by=ID), inner join by default
data_all <- merge(tidy_pd, data_cov, by = "ID")
```

## Re-arrange the data
### Reorder variables
```{r reoider variables}
# check
names(data_all)
# reorder ID, TIME, CD4, BCD4, BVL, WT, HT, BMI, CBMI, SEX, TB, using dplyr
data_all <- data_all %>% 
  select(ID, TIME, CD4, BCD4, BVL, WT, HT, BMI, CBMI, SEX, TB)
```

### Reorder rows
```{r reorder rows}
# order() the dataframe[ , ], first increasing ID and then TIME
data_all <- data_all[order(data_all$ID, data_all$TIME), ]
```

# Numerical and graphical data summary
## Numerical summary of time-invariant variables
```{r}
# use duplicated() to extract the first record of each individual
data_first <- data_all[!duplicated(data_all$ID), ]

# export data_first for further practice
write.csv(data_first, "data_first_time_invariant.csv", row.names = FALSE)

# summarize all continuous time-invariant variables BCD4, BVL, WT, HT, BMI in data_first
# summary() in basic R or dplyr
# dplyr
data_first %>% 
  select(BCD4, BVL, WT, HT, BMI) %>% 
  summary()

# basic R
# summary(data_first[ ,c("BCD4","BVL","WT","HT","BMI") ])

# summarize all categorical time-invariant variables CBMI, SEX, TB in data_first (dplyr)
data_first %>% 
  select(CBMI, SEX, TB) %>% 
  summary()

```

## Graphical summary of time-invariant variables
### Histograms
continuous time-invariant variables (BCD4, BVL, WT, HT, BMI) using for() loop
1. basic R
```{r basic R hist}
for (var in c("BCD4", "BVL", "WT", "HT", "BMI")){
  hist(data_first[[var]], # data
       main = paste("Histogram of", var),
       xlab = var,
       col = "lightblue",
       border = "white")
}

```
2. ggplot2
```{r ggplot2 hist}

```


### Correlation matrix
```{r}

```

### Box plots
```{r}

```

## Numerical summary of time-variant variables
```{r}

```

## Graphical summary of time-variant variables
### Scatterplots
#### Spaghetti plots
```{r}

```

#### Panel plots
```{r}

```

# Statistical testing
```{r}

```

